---
title: "Assignment 2: Coding Basics"
author: "Student Name"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on coding basics.

## Directions

1.  Rename this file `<FirstLast>_A02_CodingBasics.Rmd` (replacing `<FirstLast>` with your first and last name).
2.  Change "Student Name" on line 3 (above) with your name.
3.  Work through the steps, **creating code and output** that fulfill each instruction.
4.  Be sure to **answer the questions** in this assignment document.
5.  When you have completed the assignment, **Knit** the text and code into a single PDF file.
6.  After Knitting, submit the completed exercise (PDF file) to Sakai.

## Basics, Part 1

1.  Generate a sequence of numbers from one to 30, increasing by threes. Assign this sequence a name.

2.  Compute the mean and median of this sequence.

3.  Ask R to determine whether the mean is greater than the median.

4.  Insert comments in your code to describe what you are doing.

```{r basics 1}
#1. Create sequence from 1 to 30 by increments of 3. 
seq1to30 <- seq(1, 30, 3)

#2. Assign sequence's mean and median values to variables.
m <- mean(seq1to30)
md <- median(seq1to30)

#3.Determine whether sequence mean is greater than median.
m > md

```

## Basics, Part 2

5.  Create a series of vectors, each with four components, consisting of (a) names of students, (b) test scores out of a total 100 points, and (c) whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50.

6.  Label each vector with a comment on what type of vector it is.

7.  Combine each of the vectors into a data frame. Assign the data frame an informative name.

8.  Label the columns of your data frame with informative titles.

```{r basics 2}
# Create vectors with four components
students <- c("Ariel", "Amie", "Geoff", "Andrea") #string
scores <- seq(45, 60, 5) #integer
pass <- scores > 49 #logic

# Combine into dataframe
testResults <- data.frame("name" = students,
                          "score" = scores,
                          "pass" = pass)

# Print results.
testResults

```

9.  QUESTION: How is this data frame different from a matrix?

> Answer: Matrices can only contain one type of data, i.e. integer, while dataframes can contain multiple.

10. Create a function with an if/else statement. Your function should take a **vector** of test scores and print (not return) whether a given test score is a passing grade of 50 or above (TRUE or FALSE). You will need to choose either the `if` and `else` statements or the `ifelse` statement. 

11. Apply your function to the vector with test scores that you created in number 5.

```{r basics 3}

# passFunct <- function(x){
#   if(x >= 50){
#     print("TRUE")
#   }
#   else {print("FALSE")}
# }

passFunct2 <- function(x){
  ifelse(x >=50, print("TRUE"), print("FALSE"))
}

# identify whether students passed
passFunct2(testResults[,"score"])

```

12. QUESTION: Which option of `if` and `else` vs. `ifelse` worked? Why?

> Answer: "if" produces a printed output before the "else" function is used, creating asynchronous outputs that require a for loop to process more than one value.
